---
title: "X2 Missing, No Interaction, Correlated Simulation Complete Case, MICE"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "April 20, 2016"
output: pdf_document
---

### Overview

Assuming $X_1$ quantitative, $X_2$ dichotomous
$X_1|X_2$ normally distributed with mean $X_2$ and $\sigma = 1$. 


```{r,include=FALSE}
require(mosaic)
require(mice)
require(xtable)
options(xtable.comment=FALSE)
```

#### Setting up the simulation
```{r setup}
#set the number of numeration & sample size
numsim <- 250
n <- 200

#true coefficients, mse, set proportion of data missing
#probabilities for x3: .5, .3, .1, .1
beta0 <- 0.5; beta1 <- 1.5; beta2 <- -1
mse <- 2
propmiss <- .3
probx1 <- .4

```
#### Creating results vectors
```{r regparameters}
#parameters for original non-missing data
trueint <- rep(NA, numsim)
truex1 <- rep(NA,numsim) 
truex2 <- rep(NA, numsim) 

msetrueint <- rep(NA, numsim)
relbiastrueint <- rep(NA, numsim)
msetruex1 <- rep(NA, numsim)
relbiastruex1 <- rep(NA, numsim)
msetruex2 <- rep(NA, numsim)
relbiastruex2 <- rep(NA, numsim)

#parameters for complete case
ccint <- rep(NA, numsim)
ccx1 <- rep(NA,numsim) 
ccx2 <- rep(NA, numsim) 

mseccint <- rep(NA, numsim)
relbiasccint <- rep(NA, numsim)
mseccx1 <- rep(NA, numsim)
relbiasccx1 <- rep(NA, numsim)
mseccx2 <- rep(NA, numsim)
relbiasccx2 <- rep(NA, numsim)

#parameters for mice1 (default)
mice1int <- rep(NA, numsim)
mice1x1 <- rep(NA,numsim) 
mice1x2 <- rep(NA, numsim) 

msemice1int <- rep(NA, numsim)
relbiasmice1int <- rep(NA, numsim)
msemice1x1 <- rep(NA, numsim)
relbiasmice1x1 <- rep(NA, numsim)
msemice1x2 <- rep(NA, numsim)
relbiasmice1x2 <- rep(NA, numsim)

```
#### Storing coefficients
These results are not included in PDF, but shown in Rmd file.
```{r coefforloop, include=FALSE}

set.seed(202)
simforloop <- for (i in 1:numsim) {
  #generate complete dataset
  x2 <- rbinom(n, 1, prob=probx1)
  x1 <- rnorm(n,mean=x2, 1)
  
  #store true coefficients
  y <- beta0 + beta1*x1 + beta2*x2 + rnorm(n, 0, mse) 
  mod <- lm(y ~ x1 + x2) 

  coefs <- coef(mod)
  trueint[i] <- coefs[1]
  truex1[i] <- coefs[2]
  truex2[i] <- coefs[3]
  
  relbiastrueint[i] <- trueint[i] - beta0
  relbiastruex1[i] <- truex1[i] - beta1
  relbiastruex2[i] <- truex2[i] - beta2
  
  # impose missingness
  unif <- runif(n)
  cutoff <- x2*propmiss + propmiss/2  # P(missing = propmiss/2 for x2=0, 3*promiss/2 for x2=1)
  missind <- unif < cutoff
  x1miss <- x1
  x1miss[missind==0]=NA # adds missingness to x1
  ds <- data.frame(y,x1miss,x2)

  #fit cc
  cc <- lm(y ~ x1miss + x2, na.action=na.omit, data=ds)

  cccoefs <- coef(cc)
  ccint[i] <- cccoefs[1]
  ccx1[i] <- cccoefs[2]
  ccx2[i] <- cccoefs[3]
  
  relbiasccint[i] <- ccint[i] - beta0
  relbiasccx1[i] <- ccx1[i] - beta1
  relbiasccx2[i] <- ccx2[i] - beta2
  

   
  #fit mice1 (default, logistic regression)
  imp <- mice(ds, m=15, maxit=15,seed=NA)
  mice1results <- summary(pool(glm.mids(y~ x1miss + x2,family=gaussian, data=imp)))
  mice1int[i] <- mice1results[1,1]
  mice1x1[i] <- mice1results[2,1]
  mice1x2[i] <- mice1results[3,1]
  
  relbiasmice1int[i] <- mice1int[i] - beta0
  relbiasmice1x1[i] <- mice1x1[i] - beta1
  relbiasmice1x2[i] <- mice1x2[i] - beta2
}
```

#Calculating MSE

```{r}
msetrueint <- var(trueint) + relbiastrueint^2
msetruex1 <- var(truex1) + relbiastruex1^2
msetruex2 <- var(truex2) + relbiastruex2^2

mseccint <- var(ccint) + relbiasccint^2
mseccx1 <- var(ccx1) + relbiasccx1^2
mseccx2 <- var(ccx2) + relbiasccx2^2

msemice1int <- var(mice1int) + relbiasmice1int^2
msemice1x1 <- var(mice1x1) + relbiasmice1x1^2
msemice1x2 <- var(mice1x2) + relbiasmice1x2^2
```


# Saving Coefficients

Not included in PDF, but shown in Rmd file.
```{r storecoef, include=FALSE}
#true case

mtrue <- matrix(c(mean(trueint), sd(trueint),
mean(truex1), sd(truex1),
mean(truex2), sd(truex2)),byrow=TRUE, nrow=3,ncol=2)
colnames(mtrue) <- c("Mean", "Standard Deviation")
rownames(mtrue) <- c("Intercept", "x1", "x2")

mtrue

save(trueint, truex1, truex2, mtrue, 
     file="RESULTS-true.Rda")


#complete case
mcc <- matrix(c(mean(ccint), sd(ccint), 
mean(ccx1), sd(ccx1),
mean(ccx2), sd(ccx2)), byrow=TRUE, nrow=3, ncol=2)
colnames(mcc) <- c("Mean", "Standard Deviation")
rownames(mcc) <- c("Intercept", "x1", "x2")

mcc

save(ccint, ccx1, ccx2, mcc, file="RESULTS-cc.Rda")


#mice1 (default)
mmice1 <- matrix(c(mean(mice1int), sd(mice1int), 
mean(mice1x1), sd(mice1x1),
mean(mice1x2), sd(mice1x2)), byrow=TRUE, nrow=3, ncol=2)
colnames(mmice1) <- c("Mean", "Standard Deviation")
rownames(mmice1) <- c("Intercept", "x1", "x2")

mmice1


save(mice1int, mice1x1, mice1x2, mmice1,
     file="RESULTS-mice.Rda")

```

# Bias and MSE
```{r storebiasmse, include=FALSE}
#intercept
mintbias <- matrix(c(mean(relbiastrueint), mean(msetrueint),
  mean(relbiasccint), mean(mseccint),
  mean(relbiasmice1int), mean(msemice1int)), byrow=TRUE, nrow=3, ncol=2)
colnames(mintbias) <- c("Relative Bias", "MSE")
rownames(mintbias) <- c("True", "CC", "MICE")

mintbias

#x1
mx1bias <- matrix(c(mean(relbiastruex1), mean(msetruex1),
  mean(relbiasccx1), mean(mseccx1), 
  mean(relbiasmice1x1), mean(msemice1x1)), byrow=TRUE, nrow=3, ncol=2)
colnames(mx1bias) <- c("Relative Bias", "MSE")
rownames(mx1bias) <- c("True", "CC", "MICE")

mx1bias

#x2
mx2bias <- matrix(c(mean(relbiastruex2), mean(msetruex2), 
  mean(relbiasccx2), mean(mseccx2),
  mean(relbiasmice1x2), mean(msemice1x2)), byrow=TRUE, nrow=3, ncol=2)
colnames(mx2bias) <- c("Relative Bias", "MSE")
rownames(mx2bias) <- c("True", "CC", "MICE")

mx2bias




save(mintbias,mx1bias, mx2bias, file="RESULTS-biasmse.Rda")
```


```{r biasvisual}
mtrue
mcc
mmice1

mintbias
mx1bias
mx2bias
```

#### Simulation study results for one iteration
```{r truetable, results="asis"}
xmod <- xtable(mod, caption = "Truth (no missingness)", digits=3)
xmod
```

```{r cctable,results="asis", warning=FALSE}
xcc <- xtable(cc, caption = "CC", digits=3)
xcc
```


```{r mice1table,results="asis"}
xmice1 <- xtable(mice1results, caption = "Default MICE", digits=3)
xmice1
```
