---
title: "(STAT495) SQL and R: a gentle introduction (activity at Phoenix airport)"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "October 20, 2015"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


### Accessing the data from a server

```{r,message=FALSE}
require(mosaic)
options(digits=3)
trellis.par.set(theme=col.mosaic())

require(RMySQL)
con <- dbConnect(MySQL(), user="mth292", host="scidb.smith.edu", password="RememberPi", dbname="airlines")
# con = dbConnect(MySQL(), user="math230", host="mysql-research.amherst.edu", password="math230pass", dbname="math230")
```

### Descriptive statistics

```{r count}
dbGetQuery(con, "SELECT COUNT(*) as numFlights FROM flights")
```
How many flights are in the database?

SOLUTION:

### Get more of a feel for the data and SQL operations
```{r allvar}
dbGetQuery(con, "SELECT * FROM flights LIMIT 10")
```

```{r certainvar}
dbGetQuery(con, "SELECT dep_time, dep_delay, arr_time, arr_delay, carrier, origin, dest FROM flights LIMIT 10")
```

#### Let's take a look at Phoenix (PHX)
```{r phx}
ds <- dbGetQuery(con, "SELECT year, COUNT(*) as numFlights FROM flights WHERE (dest='PHX' OR origin='PHX') GROUP BY year")
names(ds)
xyplot(numFlights ~ year, type="l", data=ds)
```

What do you observe about the change in the numbers of flights to Phoenix over time?  

SOLUTION:

#### Has this changed by season?

```{r threeyear}
ds <- dbGetQuery(con, "SELECT year, month, COUNT(*) as numFlights FROM flights WHERE ((dest='PHX' OR origin='PHX') AND (year=1990 OR year=2000 OR year=2012)) GROUP BY month, year")
names(ds)
xyplot(numFlights ~ month, groups=year, auto.key = list(line = TRUE, columns=3), type=c("p", "l"), main="Flights to and from PHX by month", data=ds)
```

What do you observe about the seasonality of numbers of flights to Phoenix (and how that has changed over time)?  

SOLUTION:

#### Let's start to calculate some average arrival delays

I assert that these shouldn't include negative values.

```{r avgdelay}
dbGetQuery(con, "SELECT arr_delay, CASE WHEN (arr_delay < 0) then 0 else arr_delay end AS newdelay from flights WHERE (dest='PHX' OR origin='PHX') LIMIT 10")
```

#### What flights were there from PHX to SFO on 10/8/2012?
```{r phxsfo}
ds <- dbGetQuery(con, "SELECT * FROM flights WHERE (origin='PHX' and dest='SFO' AND year=2012 AND month=10 AND day=8)")
ds
ds = dbGetQuery(con, "SELECT dep_time, dep_delay, arr_time, arr_delay, carrier, origin, dest FROM flights WHERE (Origin='PHX' and Dest='SFO' AND year=2012 AND month=10 AND day=8)")
ds
densityplot(~ dep_time | carrier, data=ds)
```
```{r carriers}
ds = dbGetQuery(con, "SELECT * FROM carriers WHERE (name='WN' OR name='OO' or name='HP')")
ds
```
What carriers service this route?  How often and when?

SOLUTION:



#### Calculate the average flight arrival time delay, by airline
```{r warning=FALSE}
ds = dbGetQuery(con, "SELECT sum(1) as N, carrier, year, month, day, dest, avg(if(arr_delay< 0, 0, arr_delay)) as AvgArrivalDelay FROM flights WHERE (origin='PHX' and dest='SFO' AND year=2012 AND month=10 AND day=8) GROUP BY carrier")
ds
```
Verify the calculation for SkyWest Airlines (UniqueCarrier 'OO')

SOLUTION:

#### Calculate the average delay per carrier on this route for every month in 2012

Calculate the average flight arrival time delay, by airline
```{r warning=FALSE}
ds = dbGetQuery(con, "SELECT carrier, year, month, dest, avg(CASE WHEN (arr_delay < 0) then 0 else arr_delay end) as AvgArrivalDelay FROM flights WHERE (origin='PHX' and dest='SFO' AND year=2012) GROUP BY year,month,carrier")
```

```{r}
favstats(~ AvgArrivalDelay, data=ds)
densityplot(~AvgArrivalDelay, data=ds)
favstats(AvgArrivalDelay ~ month, data=ds)
xyplot(AvgArrivalDelay ~ month, groups=carrier, type=c("l"), auto.key=TRUE,data=ds)
bwplot(AvgArrivalDelay ~ as.factor(month), data=ds)
```

Describe the distribution of average delay per month by airline.  Are there carrier effects?  Are there seasonal effects?

SOLUTION:

#### Compare this by month for 1998 and 2012

Calculate the average flight arrival time delay, by airline
```{r warning=FALSE}
ds = dbGetQuery(con, "SELECT carrier, year, month, dest, avg(CASE WHEN (arr_delay < 0) then 0 else arr_delay end) as AvgArrivalDelay FROM flights WHERE (origin='PHX' and dest='SFO') AND (year=2012 OR year=1998) GROUP BY year,month,carrier")
head(ds)
```

```{r}
xyplot(AvgArrivalDelay ~ month | as.factor(year), groups=carrier, type=c("l"), auto.key=TRUE,data=ds)
bwplot(AvgArrivalDelay ~ as.factor(month) | as.factor(year), data=ds)
```

What changes do you observe over time?

SOLUTION:

### Open ended activity

Now, with your neighbor, try to identify a question of interest with these data and start to craft the SQL code to create a dataset to explore it.  Note that some queries will be *slow* the first time you run them, so think about restricting to time and airport to start.

Recall that you can run individual chunks using the *Chunks* menu which is located at the top right of the top left panel.  Be sure to run the first chunk to load mosaic and establish a connection to the server.

If you do rerun the markdown file, I'd suggest setting the `phx` chunk to be `eval=FALSE` to speed your operations.

```{r}
# your new commands go here
```

