---
title: "STAT495 (Advanced Data Analysis)"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "October 1, 2015"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

```{r, include=FALSE}
# Load additional packages here.  Uncomment the line below to use Project MOSAIC data sets.
# require(mosaicData)   
```

#### Warmup

Suppose that $Y_1$ and $Y_2$ are uniformly distributed over the triangle defined by the points (1,0), (0, 1), and (-1, 0).  Using R, find P($Y_1 \leq 3/4, Y_2 \leq 3/4)$ and P($Y_1 - Y_2 \geq 0$).  

SOLUTION:

```




```

#### Simulation 1: find the power for milestone 1 

Wild et al say that you should *make the call* if the median for one of the samples lies outside the box for the other sample.

For a sample size of 20 observations from a normal distribution in each of the two groups, find the $\alpha$ level of this test.


```{r}
require(mosaic)
require(parallel)
numsim <- 500  # number of simulations
testmilestone1 <- function(n=20, delta=0) {
  y1 <- rnorm(n, mean=0, sd=1)
  y2 <- rnorm(n, mean=delta, sd=1)
  fivenum1 <- fivenum(y1); fivenum2 <- fivenum(y2)
  if (fivenum1[3] > fivenum1[4]) {
    result <- ifelse(fivenum1[3] > fivenum2[4], "make call", "no call")
  } else {
    result <- ifelse(fivenum2[3] > fivenum1[4], "make call", "no call")
  }
  return(result=result)
}
res <- do(numsim, parallel=TRUE)*testmilestone1()
# tally(~ result, format="proportion", data=res)
```

Modify this code in the following ways to find the alpha level:

1. Group 1: generate data from an exponential distribution rather than a normal
2. Group 2: generate data from a uniform distribution rather than a normal
3. Group 3: generate data from a normal but with n=50

SOLUTION:

```{r}


```

Modify this code in the following ways to find the power level:

1. Group 1: set delta to 0.1
2. Group 2: set delta to 0.1 but change the sample size to n=50
3. Group 3: set delta to 0.1 but change the sample size to n=100

```{r}


```


#### Simulation 2: missing data

Let's consider a simple setting where $X_1$ is sometimes missing.
```{r}
set.seed(1982)
n <- 1000
x1 <- sample(0:1, n, replace=TRUE)
x2 <- rnorm(n, mean=x1, sd=1)
beta0 <- 2; beta1 <- 0.5; beta2 <- 1.5
runsim <- function(missprop=0.2) {
  y <- beta0 + beta1*x1 + beta2*x2 + rnorm(n, mean=0, sd=1)
  uniform <- runif(n, min=0, max=1)
  obsy <- ifelse(uniform < missprop, NA, y)
  truemod <- lm(y ~ x1 + x2)
  ccmod <- lm(obsy ~ x1 + x2)
  truebeta0 <- coef(truemod)[1]; truebeta1 <- coef(truemod)[2]; truebeta2 <- coef(truemod)[3]
  ccbeta0 <- coef(ccmod)[1]; ccbeta1 <- coef(ccmod)[2]; ccbeta2 <- coef(ccmod)[3]
  return(data.frame(truebeta0, truebeta1, truebeta2, ccbeta0, ccbeta1, ccbeta2))
}
```

```{r}
set.rseed(1982)
res <- do(20, parallel=TRUE)*runsim()    # change this to a larger value!
head(res)
```

ASSIGNMENT: calculate the relative efficiency of the complete case estimator relative to the true model as a function of the proportion missing.



